
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


Le notebook qui suit porte sur le projet de 4ème année en Machine Learning. Le but est d'améliorer la prévision de l'émission en CO2. Le jeu de données provient d'un concours Kaggle, nommé "global-data-on-sustainable-energy.csv"

On importe nos librairies ainsi que notre jeu de données :

```{r}
library(plotly)
library(dplyr)
library(corrplot)
library(ggplot2)
library(factoextra)
library(FactoMineR)
library(gridExtra)
library(ggbiplot)
library(tidyverse)
library(GGally)
library(reshape2)
library(glmnet) 
library(ggfortify)
library(pROC)
library(ROCR)
```

#*Part I : Exploratory data analysis*
##*Question 1*
- Start by checking the nature of the different variables and their encoding. 
- Convert the variable Year into a qualitative variable.
- Convert the variable Density (P/Km2) into a numeric variable. 

```{r}
#Load and print the 5 first lines of the DataFrame
data <- read.csv("global-data-on-sustainable-energy (1).csv", header = TRUE)
head(data)
```

```{r}
#Print the type of each variables of the DataFrame
str(data)
```

```{r}
#Print different characteristic of the different variables in the DataFrame
summary(data)
```

```{r}
#Transform the variable 'Year' in a qualitative variable
data$Year <- as.factor(data$Year)
class(data$Year)
```

```{r}
#Transform the variable 'Density' in a numeric variable
data$Density.n.P.Km2.<-as.numeric(gsub(",","",data$Density.n.P.Km2.))
class(data$Density.n.P.Km2.)
```


```{r}
#Verify that we have the right types for the variables 'Years' and 'Density' 
str(data)
```
##*Question 2*
- Determine the rate of missing values for each variable.
- For this project, we propose to remove the variables with a very high rate of missing data:
    - Renewable-electricity-generating-capacity-per-capita
    - Financial flows to developing countries (US Dollars) 
    - Renewables (% equivalent primary energy)

```{r}
#Print the number of missing value for each vraiables in the DataFrame
missing_values <- colSums(is.na(data))
print(missing_values)
```

```{r}
#Creating a bar plot showing these missing values 
fig <- plot_ly(x = names(missing_values), y = missing_values, type = "bar", 
               name = "Missing values for each variables", 
               marker = list(color = "purple"),
               xlab = "Variables", ylab = "Number of missing values",
               title = "Missing values for each variables")

fig
```

```{r}
#Retrieving the variable's full names
names(data)
```

```{r}
#Deleting the columns corresponding tp the variables with too many missing values
data <- subset(data, select = -c(Renewable.electricity.generating.capacity.per.capita, Financial.flows.to.developing.countries..US..., Renewables....equivalent.primary.energy.))

```


```{r}
#Verifying that the wanted variables are well deleted from the DataFrame
names(data)
```

##*Question 3*
For the rest of the study, you will create a dataset containing only those individuals with no missing
values. This will leave 2868 observations.

```{r}
# Creating a new DataFrame containing only individuals with no missing values
data_full <- data[complete.cases(data), ]
```


```{r}
# Verifying that we have 2868 observations in the complete dataset
nrow(data_full)
str(data_full)
```
##*Question 4*
Start your exploration with a unidimensional descriptive analysis of the data. 
Do you think transformations of quantitative variables are relevant?

```{r}
ggplot(data_full, aes(y = `Access.to.electricity....of.population.`)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Boxplot for Access-to-electricity variable") +
  scale_x_continuous(limits =c(-2,2))
```

```{r}
# Other type of descriptive analysis of qualitative variable
ggplot(data_full, aes(x = `Electricity.from.fossil.fuels..TWh.`)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "black") +
  labs(title = "Electricity-from-fossil-fuels",
       x = "Electricity-from-fossil-fuels (TWh)",
       y = "Fréquence") +
  theme_minimal()

# We note that this variable has a lot of data between 0 and 500 TWh
```
```{r}
ggplot(data_full, aes(x = `Electricity.from.fossil.fuels..TWh.`)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(title = "Densité de Electricity-from-fossil-fuels")
```


```{r}
ggplot(data_full, aes(x = `Access.to.clean.fuels.for.cooking`)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Densité de Access.to.clean.fuels.for.cooking") + 
  scale_x_continuous(limits =c(-25,125))
```

```{r}
# Drawing the density graph
ggplot(data_full, aes(x=`Primary.energy.consumption.per.capita..kWh.person.`)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "The density of Primary energy consumption per capita (kWh/person)") + 
  scale_x_continuous(limits =c(-25000,200000))
```

```{r}
#We apply a logarithmic transformation to this variable
data_full$Primary_energy_consumption_per_capita_log <- log(data_full$Primary.energy.consumption.per.capita..kWh.person.)

#Deleting the variable without the logarithmic transformation
data_full <- subset(data_full, select = -c(Primary.energy.consumption.per.capita..kWh.person.))
str(data_full)
```


```{r}
#Displaying the new density of Primary Energy Consumption Per Capita (kWh/person)"
ggplot(data_full, aes(x=`Primary_energy_consumption_per_capita_log`)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "The density of Primary energy consumption per capita log (kWh/person)")

#We have performed our logarithmic transformation on the "Primary.energy.consumption.per.capita..kWh.person." variable.
```
##*Question 5*
Visualize the great heterogeneity of CO2 emissions between countries. 
Which 5 countries emit the most CO2?

```{r}
#Renaming the "Value_co2_emissions_kt_by_country_new" column to "CO2". 
data_full <- data_full %>%
  rename(CO2 = Value_co2_emissions_kt_by_country)
```

```{r}
#We verify :
names(data_full)
```


```{r}
#We can observe the emission of CO2 for each country

ggplot(data_full, aes(x = Entity, y = CO2)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Emission of CO2 for each country",
       x = "Country",
       y = "Emission of CO2 (in Kilo Tonnes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

#The legend is difficult to read, but there are differences between countries.
```

```{r}
#Now we want to see the 5 countries with the highest CO2 emissions:


#Computing the maximum CO2 emissions by country
CO2_max <- data_full %>%
  group_by(Entity) %>%
  summarise(CO2 = max(CO2)) %>%
  arrange(desc(CO2)) %>%
  head(5)

ggplot(CO2_max, aes(x = reorder(Entity, CO2), y = CO2, fill = CO2)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 5 countries with the highest CO2 emissions",
       x = "Country",
       y = "Emissions of CO2 in KiloTonnes",
       fill = "Emissions de CO2") +
  scale_fill_gradient(low = "blue", high = "yellow") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotation des noms de pays sur l'axe x for better visibility
```
##*Question6*
Continue with a multidimensional descriptive analysis. Use visualization techniques: e.g. scatterplots, correlation graphs... 
Analyze dependencies between quantitative variables.
```{r}
str(data_full)
names(data_full)
```


```{r}
# We rename the columns of the Full DataFrame
data2 <- data_full %>%
  rename(Access_to_electricity = Access.to.electricity....of.population.,
         Access_to_clean_fuels_for_cooking = Access.to.clean.fuels.for.cooking,
         Renewable_energy_share_in_the_total_final_energy_consumption = Renewable.energy.share.in.the.total.final.energy.consumption....,
         Electricity_from_fossil_fuels = Electricity.from.fossil.fuels..TWh.,
         Electricity_from_nuclear = Electricity.from.nuclear..TWh.,
         Electricity_from_renewables = Electricity.from.renewables..TWh.,
         Low_carbon_electricity = Low.carbon.electricity....electricity.,
         Energy_intensity_level_of_primary_energy = Energy.intensity.level.of.primary.energy..MJ..2017.PPP.GDP.,
         Density_n_P_Km2 = Density.n.P.Km2.,
         Land_Area_Km2 = Land.Area.Km2.)

names(data2)
```

```{r}
# Convert the "Land_Area-Km2" variable from integer to real
data2 <- data2 %>%
  mutate(Land_Area_Km2 = as.numeric(Land_Area_Km2))
str(data2)
```


```{r}
cols <- c('Access_to_electricity',
          'Access_to_clean_fuels_for_cooking',
          'Renewable_energy_share_in_the_total_final_energy_consumption',
          'Electricity_from_fossil_fuels',
          'Electricity_from_nuclear',
          'Electricity_from_renewables',
          'Low_carbon_electricity',
          'Energy_intensity_level_of_primary_energy',
          'CO2',
          'gdp_growth',
          'gdp_per_capita',
          'Density_n_P_Km2',
          'Land_Area_Km2',
          'Latitude',
          'Longitude',
          'Primary_energy_consumption_per_capita_log')


correlation_matrix <- cor(data2[cols])


corrplot(correlation_matrix, 
          method = "color", 
          type = "upper", 
          tl.col = "black", 
          tl.srt = 10, 
          tl.cex = 0.7,
          diag = FALSE,  
          addCoef.col = "black",  
          number.cex = 0.4,  
          tl.pos = "lt")
```
```{r}
corrplot(correlation_matrix, 
          method = "ellipse",  # Ellipse Method instead of the numeric one
          tl.pos = "lt", type = "upper", 
          tl.col = "black", 
          tl.srt = 10, 
          tl.cex = 0.6)

```


```{r}
#Plot a scatterplot for all correlation variables

pairs(data2[, c('Access_to_electricity', 'Access_to_clean_fuels_for_cooking', 'Renewable_energy_share_in_the_total_final_energy_consumption', 'Electricity_from_fossil_fuels', 'Electricity_from_nuclear', 'Electricity_from_renewables', 'Low_carbon_electricity', 'Primary_energy_consumption_per_capita_log', 'Energy_intensity_level_of_primary_energy', 'CO2', 'gdp_growth', 'gdp_per_capita', 'Density_n_P_Km2', 'Land_Area_Km2', 'Latitude', 'Longitude')], main = "Scatterplot pour les variables de corrélation")

#We can't see much, so we'll try to draw a scatterplot for the two most correlated variables.

```

```{r}
#We want to see which pairs of variables are most correlated
paires_plus_cor <- which(correlation_matrix == max(correlation_matrix[lower.tri(correlation_matrix)], na.rm = TRUE), arr.ind = TRUE)

#Displaying these variables
var_plus_cor <- colnames(correlation_matrix)[paires_plus_cor[, 1]]
var_plus_cor
```

```{r}
#A scatterplot is displayed to show the relationships between the most correlated pairs of variables.
scatterplot <- ggplot(data2, aes(x = Electricity_from_fossil_fuels, y = CO2)) +
  geom_point() +  
  labs(title = "Relation entre les variables les plus corrélées : C02 et Electricity_from_fossil_fuels",
       x = "Electricity_from_fossil_fuels",
       y = "CO2")

scatterplot
```

##*Question 7*
Perform a principal component analysis of quantitative variables and interpret the results.
```{r}
#We create a new dataframe to keep only quantitative variables
data_quanti <- subset(data2, select = -c(Entity, Year))

#PCA is performed
#Set to "true" to standardize the data

pca <- PCA(data2,quali.sup=c(1,2), scale. = TRUE,graph=FALSE)
summary(pca)

fviz_pca_ind(pca,geom=c("point"), habillage = data2[,2])

```
In 2 dimensions, 50% of the variance can be explained, and many outliers can be identified.

```{r}
#Printing how well the variables are represented, so printing how well the axes explain the variables
q1=fviz_pca_var(pca, col.var="cos2",repel=TRUE,gradient.cols=c("#00afbb","red","yellow"))
q1
```

```{r}
#Displays the contribution of variables to the axes
c1=fviz_pca_var(pca,col.var="contrib",repel=TRUE,gradient.cols=c("#00afbb","red","yellow"))
c1
```

```{r}
q2=fviz_pca_ind(pca, col.ind="cos2",geom=c("point"),gradient.cols=c("#00afbb","red","yellow"))
q2
```
Individuals close to 0 are poorly explained

```{r}
c2=fviz_pca_ind(pca, col.ind="contrib",geom=c("point"),gradient.cols=c("#00afbb","red","yellow"))
c2
```

```{r}

#We keep the eigenvalues of the PCA object, which correspond to the amount of variance explained by each principal component.
variance_explained <- pca$sdev^2

#The proportion of variance explained is calculated: the contribution of each principal component to the total variance of the data.
proportion_var <- variance_explained / sum(variance_explained)

#The variance explained by the principal components is displayed.
data_var <- data.frame(Composante = 1:length(var_explained), Variance_Explained = proportion_var)

ggplot(data_var, aes(x = Composante, y = Variance_Explained)) +
  geom_line(color = "purple") +
  geom_point(color = "purple") +
  labs(title = "Variance explained by each principal component",
       x = "Principales Components",
       y = "Proportion of variance explained") +
  theme_minimal()
```


```{r}
#The higher the proportion of variance explained for a given principal component, the more important that principal component is in representing the structure of the data: the 1st component is the most important here.
```

##*Question 8*
Visualize the possible dependency between the variable Year and the variable to be predicted.

```{r}
#Here we need to work with the DataFrame "data2" to obtain the qualitative variables

#A linear dependency is displayed between the variables C02 and Year

ggplot(data2, aes(x = Year, y = CO2)) +
  geom_point() +  
 labs(title = "Dependency between year and CO2 emissions",
       x = "Year",
       y = "CO2 Emissions (in KiloTonnes)")
```

```{r}
#We do the same thing but with an interactive graph to get the exact CO2 value at each point.
plot_ly(data = data2, x = ~Year, y = ~CO2, type = "scatter", mode = "markers")
```


```{r}
#A violin graph of CO2 emissions per year
ggplot(data2, aes(x = Year, y = CO2)) +
  geom_violin(fill = "purple") +
  labs(title = "Yearly Emission of CO2",
       x = "Year",
       y = "Emissions of CO2 (in KiloTonnes)")
```

```{r}
#Box diagram of CO2 emissions per year

ggplot(data2, aes(x = Year, y = CO2)) +
  geom_boxplot(fill = "purple") +
  labs(title = "#Box diagram of CO2 emissions per year",
       x = "Year",
       y = "Emissions of CO2 (in KiloTonnes)")

```



```{r}
#We create a DataFrame with the yearly maximum values of CO2

CO2_max_year <- aggregate(data2$CO2, by=list(Year=data2$Year), FUN=max)
CO2_max_year <- CO2_max_year[order(CO2_max_year$x),]
CO2_max_year20 <- head(CO2_max_year, 20)
names(CO2_max_year20) <- c("Year", "CO2")

# Creating a barplot
fig <- ggplot(CO2_max_year20, aes(x = Year, y = CO2, fill = CO2)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Emissions of CO2", title = "Maximum amount of CO2 per year") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45))

fig
```

```{r}
#The CO2 maximum is unwise because we have a lot of outliers.
#We create a DataFrame with the average CO2 values per year

CO2_moy_year <- aggregate(data2$CO2, by=list(Year=data2$Year), FUN=mean)
CO2_moy_year <- CO2_moy_year[order(CO2_moy_year$x),]
CO2_moy_year20 <- head(CO2_moy_year, 20)
names(CO2_moy_year20) <- c("Year", "CO2")

# Creating a barplot
fig <- ggplot(CO2_moy_year20, aes(x = Year, y = CO2, fill = CO2)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Emissions of CO2", title = "Average quantity of CO2 per year") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45))

fig
```

```{r}
#Box diagram of CO2 emissions per year
data_log <- data2
data_log <- data_log[, 11, drop = FALSE]  # Keep the DataFrame structure
data_log[, 1] <- log(data_log[, 1])  # Apply log to the extracted column
data_log$Year <- data2$Year  # Reallocate the Year column which doesn't have to be transformed into a logarithmic scale


ggplot(data_log, aes(x = Year, y = CO2)) +
  geom_boxplot(fill = "purple") +
  labs(title = "Box diagram of CO2 emissions per year",
       x = "Year",
       y = "CO2 Emissions  (in KiloTonnes)")

```
```{r}
#We will use a 1-factor ANOVA to analyse the dependence of a quantitative variable on a qualitative variable.
anova1 <- lm(CO2 ~ Year, data = data2)
summary(anova1)
```


```{r}
#Creating a 2nd model, the constant one, to compare with the previous one. 
anova2 <- lm(CO2 ~ 1, data = data2)

#We compare the sub-model with the first model
anova(anova2,anova1)
#We have a p-val>0.05, so we accept this sub-model.
```
```{r}
#These results are verified using a chi2 test of independence.
chi2_test <- chisq.test(data2$CO2, data2$Year)
print(chi2_test)

#pval=0.14>0.05: do not reject H0, CO2 and Year are independent
```


```{r}
#We realise that in the tests we can remove the year, but on the graphs of average CO2 quantities we can still see an increase over the years.

#So we're trying to create a model based on the country and the year.
anova3 <- lm(CO2 ~ Year+Entity, data = data2)
summary(anova3)

#The R-quared's value is 0.93, so the model is much better. 
```

```{r}
#We make a sub-model depending only on the country
anova4=lm(CO2 ~ Entity, data = data2)
#Compare this sub-model with the previous one
anova(anova4,anova3)

#We can see that pval=0.15>0.05, so we can keep the sub-model and the effect of the year is in fact negligible compared with that of the country.
```

#*Part II : Modelisation*